---
title: "Functions"
author: "Tim Trice"
output: workflowr::wflow_html
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE, 
  include = TRUE, 
  eval = TRUE, 
  message = TRUE, 
  warning = TRUE, 
  error = FALSE
)
```

```{r libraries}
library(readr)
```

## details

```{r details, echo = FALSE}
details <- 
  read_csv(
    file = here::here("./output/details.csv"), 
    col_types = cols(
      .default = col_character(),
      EPISODE_ID = col_integer(), 
      EVENT_ID = col_integer(),
      STATE_FIPS = col_integer(),
      CZ_FIPS = col_integer(),
      BEGIN_DATE_TIME = col_datetime(format = ""),
      END_DATE_TIME = col_datetime(format = ""),
      INJURIES_DIRECT = col_integer(),
      INJURIES_INDIRECT = col_integer(),
      DEATHS_DIRECT = col_integer(),
      DEATHS_INDIRECT = col_integer(),
      DAMAGE_PROPERTY = col_number(),
      DAMAGE_CROPS = col_number(),
      MAGNITUDE = col_double(),
      TOR_LENGTH = col_double(),
      TOR_WIDTH = col_double(),
      BEGIN_RANGE = col_integer(),
      END_RANGE = col_integer(),
      BEGIN_LAT = col_double(),
      BEGIN_LON = col_double(),
      END_LAT = col_double(),
      END_LON = col_double()
    )
  )
```

### Description

### Usage

```r
details
```

### Format

```{r, echo = FALSE}
tibble::glimpse(details)
```

### Source

### Examples

```{r ref.label = "details", eval = FALSE}
```

## episode_narratives

```{r episode_narratives, echo = FALSE}
episode_narratives <- 
  read_csv(
    file = here::here("./output/episode_narratives.csv"), 
    col_types = cols()
  )
```

### Description

### Usage

```r
episode_narratives
```

### Format

```{r, echo = FALSE}
tibble::glimpse(episode_narratives)
```

### Source

### Examples

```{r ref.label = "episode_narratives"}
```

## event_narratives

```{r event_narratives, echo = FALSE}
event_narratives <- 
  read_csv(
    file = here::here("./output/event_narratives.csv"), 
    col_types = cols(
      EPISODE_ID = col_integer(),
      EVENT_ID = col_integer(),
      EVENT_NARRATIVE = col_character()
    )
  )

```

### Description

### Usage

```r
event_narratives
```

### Format

```{r, echo = FALSE}
tibble::glimpse(event_narratives)
```

### Source

### Examples

```{r ref.label = "event_narratives"}
```

## event_types

```{r event_types, echo = FALSE}
event_types <- 
  read_csv(
    file = here::here("./output/event_types.csv"), 
    col_types = cols(
      EPISODE_ID = col_number(), 
      EVENT_ID = col_number(), 
      EVENT_TYPE = col_character()
    )
  )
```

### Description

### Usage

```r
event_types
```

### Format

```{r, echo = FALSE}
tibble::glimpse(event_types)
```

### Source

### Examples

```{r ref.label = "event_types"}
```

## fatalities

```{r fatalities, echo = FALSE}
fatalities <- 
  read_csv(
    file = here::here("./output/fatalities.csv"), 
    col_types = cols()
  )
```

### Description

### Usage

```r
fatalities
```

### Format

```{r, echo = FALSE}
tibble::glimpse(fatalities)
```

### Source

### Examples

```{r ref.label = "fatalities"}
```

## fips

```{r fips, echo = FALSE}
fips <- 
  read_csv(
    file = here::here("./output/fips.csv"), 
    col_types = cols()
  )
```

### Description

### Usage

```r
fips
```

### Format

```{r, echo = FALSE}
tibble::glimpse(fips)
```

### Source

### Examples

```{r ref.label = "fips"}
```

## locations

```{r locations, echo = FALSE}
locations <- 
  read_csv(
    file = here::here("./output/locations.csv"), 
    col_types = cols()
  )
```

### Description

### Usage

```r
locations
```

### Format

```{r, echo = FALSE}
tibble::glimpse(locations)
```

### Source

### Examples

```{r ref.label = "locations"}
```

## zone_county

```{r zone_county, echo = FALSE}
zone_county <- 
  read_csv(
    file = here::here("./output/zone_county.csv"), 
    col_types = cols()
  )
```

### Description

### Usage

```r
zone_county
```

### Format

```{r, echo = FALSE}
tibble::glimpse(zone_county)
```

### Source

### Examples

```{r ref.label = "zone_county"}
```
