---
title: "Tornadoes"
author: "Tim Trice"
date: "2019-03-03"
output: workflowr::wflow_html
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE, 
  message = FALSE, 
  warning = TRUE, 
  error = TRUE, 
  fig.width = 10, 
  fig.asp = 1, 
  fig.dpi = 180
)
```

```{r libraries}
library(kableExtra)
library(tidyverse)
```

```{r data}
details <- 
  read_csv(
    file = here::here("./output/details.csv"), 
    col_types = cols(
      .default = col_character(),
      EPISODE_ID = col_integer(), 
      EVENT_ID = col_integer(),
      STATE_FIPS = col_integer(),
      CZ_FIPS = col_integer(),
      BEGIN_DATE_TIME = col_datetime(format = ""),
      END_DATE_TIME = col_datetime(format = ""),
      INJURIES_DIRECT = col_integer(),
      INJURIES_INDIRECT = col_integer(),
      DEATHS_DIRECT = col_integer(),
      DEATHS_INDIRECT = col_integer(),
      DAMAGE_PROPERTY = col_number(),
      DAMAGE_CROPS = col_number(),
      MAGNITUDE = col_double(),
      TOR_LENGTH = col_double(),
      TOR_WIDTH = col_double(),
      BEGIN_RANGE = col_integer(),
      END_RANGE = col_integer(),
      BEGIN_LAT = col_double(),
      BEGIN_LON = col_double(),
      END_LAT = col_double(),
      END_LON = col_double()
    )
  )

bad_lon <- 
  filter(details, !between(BEGIN_LON, -180, 180) | !between(END_LON, -180, 180))
```

#### Fujita Scales

```{r tornado-scales}
tornado_scales <- 
  tibble(
    EF_scale = c("EF0", "EF1", "EF2", "EF3", "EF4", "EF5"), 
    EF_wind = c("65-85", "86-110", "111-135", "136-165", "166-200", ">200"), 
    F_scale = c("F0", "F1", "F2", "F3", "F4", "F5"), 
    F_wind = c("40-72", "73-112", "113-157", "158-206", "207-260", "261-318")
  )

tornado_scales %>% 
  kable() %>% 
  kable_styling(c("striped", "bordered")) %>% 
  add_header_above(c("Enhanced Fujita Wind Scale" = 2L, "Fujita Scale" = 2L))
```

```{r}
enhanced_fujita <- 
  tibble(
    Wind = seq(65, 320, by = 1), 
    Scale = case_when(
      between(Wind, 65, 85)   ~ "EF0", 
      between(Wind, 86, 110)  ~ "EF1", 
      between(Wind, 111, 135) ~ "EF2", 
      between(Wind, 136, 165) ~ "EF3", 
      between(Wind, 166, 200) ~ "EF4", 
      between(Wind, 200, 320) ~ "EF5", 
    )
  )

fujita <- 
  tibble(
    Wind = seq(40, 318, by = 1), 
    Scale = case_when(
      between(Wind, 40, 72)   ~ "F0", 
      between(Wind, 73, 112)  ~ "F1", 
      between(Wind, 113, 157) ~ "F2", 
      between(Wind, 158, 206) ~ "F3", 
      between(Wind, 207, 260) ~ "F4", 
      between(Wind, 261, 318) ~ "F5", 
    )
  )
```

```{r plot-enhanced-fujita, fig.asp = 0.5, caption = "Enhanced Fujita Scale"}
enhanced_fujita %>% 
  ggplot() + 
  aes(x = Wind, y = Scale, color = Scale) + 
  geom_line(size = 10) + 
  scale_x_continuous(
    limits = c(0, 320), 
    breaks = seq(0, 320, by = 20), 
    minor_breaks = seq(0, 320, by = 10), 
    expand = c(0, 0)
  ) + 
  theme(legend.position = "bottom") + 
  guides(color = guide_legend(title = "Enhanced Fujita Scale", nrow = 1L))
```

```{r plot-fujita, fig.asp = 0.5, caption = "Fujita Scale"}
fujita %>% 
  ggplot() + 
  aes(x = Wind, y = Scale, color = Scale) + 
  geom_line(size = 10) + 
  scale_x_continuous(
    limits = c(0, 320), 
    breaks = seq(0, 320, by = 20), 
    minor_breaks = seq(0, 320, by = 10), 
    expand = c(0, 0)
  ) + 
  theme(legend.position = "bottom") + 
  guides(color = guide_legend(title = "Fujita Scale", nrow = 1L))
```

#### Invalid `TOR_F_SCALE` Values

```{r}
details %>%
  filter(EVENT_TYPE == "Tornado") %>% 
  #' Remove bad latitude/longitude values
  group_by(TOR_F_SCALE) %>% 
  summarise(n = n())
```

```{r}
details %>% 
  filter(
    EVENT_TYPE == "Tornado", 
    is.na(TOR_F_SCALE) | TOR_F_SCALE == "EFU"
  ) %>% 
  select(EPISODE_ID, EVENT_ID, EVENT_TYPE, TOR_F_SCALE, MAGNITUDE, MAGNITUDE_TYPE) %>% 
  group_by(TOR_F_SCALE, MAGNITUDE, MAGNITUDE_TYPE) %>% 
  summarise(n = n()) %>% 
  arrange(TOR_F_SCALE, MAGNITUDE, MAGNITUDE_TYPE, n) %>% 
  kable() %>% 
  kable_styling("striped")
```

Where `TOR_F_SCALE` is NA or equals "EFU", there is no valuable data to correct this. The variable `MAGNITUDE` is used to mark the measured extent of wind speeds (or hail, in some cases) and `MAGNITUDE_TYPE` identifies how that observation was obtained. Both of these values are NA for "EFU" observations, and `MAGNITUDE` does not reflect valid wind speed values for a tornado where `TOR_F_SCALE` is NA. 

With this, can make `TOR_F_SCALE` a `NA_character_` value where `TOR_F_SCALE` is "EFU".

```{r}
details$TOR_F_SCALE[details$TOR_F_SCALE == "EFU"] <- NA_character_
```

```{r plot-tornadoes, fig.asp = 0.72}
details %>%
  filter(
    EVENT_TYPE == "Tornado", 
    !is.na(TOR_F_SCALE)
  ) %>% 
  #' Remove observations with bad latitude/longitude values
  setdiff(bad_lon) %>% 
  select(EVENT_TYPE, TOR_F_SCALE, BEGIN_LAT, BEGIN_LON) %>% 
  extract(
    col = TOR_F_SCALE, 
    into = "SCALE", 
    regex = "^E*F([[:digit:]])$", 
    convert = TRUE, 
    remove = FALSE
  ) %>% 
  mutate(SCALE = fct_reorder(.f = as.character(SCALE), .x = SCALE)) %>% 
  ggplot() + 
  aes(
    x = BEGIN_LON, 
    y = BEGIN_LAT, 
    color = SCALE
  ) +
  geom_polygon(
    data = rnaturalearth::countries110, 
    aes(x = long, y = lat, group = group), 
    fill = "white", 
    color = "black", 
    size = 0.1
  ) + 
  geom_point() + 
  coord_cartesian() + 
  guides(color = guide_legend(nrow = 1L)) +
  viridis::scale_color_viridis(discrete = TRUE) +
  # scale_color_brewer(palette = "Set1") +
  theme_void() + 
  theme(legend.position = "bottom")
```

Where `TOR_F_SCALE` is "F*", do we have valid `MAGNITUDE` values?

```{r}
details %>% 
  filter(grepl("^F[[:digit:]]$", TOR_F_SCALE)) %>% 
  select(TOR_F_SCALE, MAGNITUDE) %>% 
  group_by(TOR_F_SCALE) %>% 
  summarise(
    Min = min(MAGNITUDE), 
    Max = max(MAGNITUDE)
  )
```

...and "EF*"?

```{r}
details %>% 
  filter(grepl("^EF[[:digit:]]$", TOR_F_SCALE)) %>% 
  select(TOR_F_SCALE, MAGNITUDE) %>% 
  group_by(TOR_F_SCALE) %>% 
  summarise(
    Min = min(MAGNITUDE), 
    Max = max(MAGNITUDE)
  )
```

So, at least one record for each scale has a NA value. 

#### Severe Tornadoes

Under the Enhanced Fujita Scale, tornadoes rated EF4 or higher (winds > 166mph) produce devastating damage. Corresponding to the Fujita Scale, this would be at least a F3. 

Plot tornadic events produced by at least a F3 or EF4 tornado.

```{r}
details %>% 
  filter(TOR_F_SCALE %in% c("F3", "F4", "F5", "EF4", "EF5")) %>% 
  setdiff(bad_lon) %>% 
  select(EVENT_TYPE, TOR_F_SCALE, BEGIN_LAT, BEGIN_LON) %>% 
  extract(
    col = TOR_F_SCALE, 
    into = "SCALE", 
    regex = "^E*F([[:digit:]])$", 
    convert = TRUE, 
    remove = FALSE
  ) %>% 
  mutate(SCALE = fct_reorder(.f = as.character(SCALE), .x = SCALE)) %>% 
  ggplot() + 
  aes(
    x = BEGIN_LON, 
    y = BEGIN_LAT, 
    color = SCALE
  ) +
  borders("state") +
  geom_point() +
  coord_map() +
  theme_void() + 
  labs(
    title = "Tornadic events of at least EF4 or F3", 
    subtitle = "No events exist outside of the continental United"
  )
```

#### Tornado Width

```{r}
details %>% 
  filter(!is.na(TOR_WIDTH), !is.na(TOR_F_SCALE)) %>% 
  select(TOR_F_SCALE, TOR_WIDTH) %>% 
  ggplot() + 
  aes(x = TOR_F_SCALE, y = TOR_WIDTH, fill = TOR_F_SCALE) + 
  geom_boxplot() + 
  coord_flip() + 
  scale_y_continuous(
    labels = scales::comma
  ) + 
  theme(legend.position = "none") + 
  labs(
    title = "Tornado Width by Scale", 
    x = "(Enhanced) Fujita Scale",
    y = "Tornado Width (feet)"
  )
```

#### Tornado Length (Distance Travelled)

```{r}
details %>% 
  filter(!is.na(TOR_LENGTH), !is.na(TOR_F_SCALE)) %>% 
  select(TOR_F_SCALE, TOR_LENGTH) %>% 
  ggplot() + 
  aes(x = TOR_F_SCALE, y = TOR_LENGTH, fill = TOR_F_SCALE) + 
  geom_boxplot() + 
  coord_flip() + 
  theme(legend.position = "none") + 
  labs(
    title = "Tornado Length by Scale", 
    x = "(Enhanced) Fujita Scale",
    y = "Tornado Length (miles)"
  )
```

```{r}
details %>% 
  filter(
    !is.na(TOR_LENGTH), 
    !is.na(TOR_F_SCALE), 
    TOR_LENGTH <= 100
  ) %>% 
  select(TOR_F_SCALE, TOR_LENGTH) %>% 
  ggplot() + 
  aes(x = TOR_F_SCALE, y = TOR_LENGTH, fill = TOR_F_SCALE) + 
  geom_boxplot() + 
  coord_flip() + 
  theme(legend.position = "none") + 
  labs(
    title = "Tornado Length by Scale", 
    subtitle = "Tornadoes with a length <= 100 miles", 
    x = "(Enhanced) Fujita Scale",
    y = "Tornado Length (miles)"
  )
```

